---
title: "Exercise2"
output: html_document
date: "2023-04-28"
---
# Preparatory Tasks
```{r}
##demonstrate lead and lag

library(dplyr)

#lead() and lag() return a vector of the same length as the input, 
#just offset by a specific number of values (default is 1). Consider the following sequence:

mynumbers <- 1:10



wildschwein <- tibble(
  TierID = c(rep("Hans", 5), rep("Klara", 5)),
  DatetimeUTC = rep(as.POSIXct("2015-01-01 00:00:00", tz = "UTC") + 0:4 * 15 * 60, 2)
)

wildschwein





#To calculate the timelag (difference) with base-R, we need to mention wildschwein three times

wildschwein$diff <- as.numeric(difftime(lead(wildschwein$DatetimeUTC),wildschwein$DatetimeUTC))

#group data frame first
wildschwein <- group_by(wildschwein,TierID)

#Using mutate() we can simplify this operation slightly:


wildschwein <- mutate(wildschwein, diff = as.numeric(difftime(lead(DatetimeUTC), DatetimeUTC)))

wildschwein

#Piping makes it easier

wildschwein <- wildschwein |> 
  group_by(TierID) |> 
  mutate(diff = as.numeric(difftime(lead(DatetimeUTC), DatetimeUTC)))

wildschwein

```

# Task 1 Import Data
```{r setup, include=FALSE}
library("readr") 
library("sf") 
library(dplyr)
library(ggplot2)
library(tmap)

wildschwein_BE <- read_delim("wildschwein_BE_2056.csv", ",")

wildschwein_BE <- st_as_sf(wildschwein_BE, coords = c("E", "N"), crs = 2056, remove = FALSE)
```

## Task 2: Getting an overview

Calculate the time difference between subsequent rows as described in the demo. You can calculate the time difference using the function difftime() in combination with lead().

the function difftime() has an option units. Set this to secs to get the time difference in seconds
use as.integer() to turn the output returned by difftime() into an integer.
store the output in a new column (e.g. timelag)
Now inspect your data in more detail. Try to answer the following questions:

- How many individuals were tracked?
- For how long were the individual tracked? Are there gaps?
- Were all individuals tracked concurrently or sequentially?
- What is the temporal sampling interval between the locations?

```{r}
wildschwein_BE$TierName |> unique()
group_by(wildschwein_BE,TierName) #there are 3 individuals tracked

ggplot(wildschwein_BE,aes(DatetimeUTC,TierName))+ #sabi has been tracked longer
  geom_point()

wildschwein$diff_s

wildschwein <- wildschwein_BE |> 
  group_by(TierName) |> 
  mutate(diff_s = as.numeric(difftime(lead(DatetimeUTC),DatetimeUTC)))


ggplot(wildschwein, aes(diff_s/60))+
  geom_histogram((binwidth =1)) +
  lims(x=c(0,5000/60))+
  scale_y_log10()


ggplot(wildschwein,aes(DatetimeUTC,TierName$Sabi))+ #sabi has been tracked longer
  geom_point()


wildschwein |> 
  filter(DatetimeUTC < "2014-08-24") |> 
  ggplot(DatetimeUTC,diff_s,colour =TierName)+
  geom_point()+
  geom_line()


#euclidian distance
wildschwein
N1 <- 1204752
N2 <- 1204863
E1 <- 2570409
E2 <- 2570402

wildschwein <- wildschwein |>
  group_by(TierName) |> 
  mutate(steplength_m = sqrt((E-lead(E))^2+ (N-lead(N))^2))

wildschwein <- wildschwein |> 
  mutate(speed_ms = steplength_m/diff_s)

hist(log10(wildschwein$speed_ms))

```


#Task 6

```{r}
library(ggmap)


posmo_neu <- st_as_sf(posmo_data, coords = c("lon_x", "lat_y"), crs = 2056, remove = FALSE)
```

```{r}
posmo_korr <- posmo_neu %>%
  mutate(transport_mode = ifelse(transport_mode == "Car", "Bike", transport_mode))

ggplot(posmo_korr,aes(datetime,transport_mode))+ #a lot of NA trackings
  geom_point()


posmo_neu <- posmo_korr |> 
  group_by(weekday,transport_mode) |>
  mutate(diff_s = as.numeric(difftime(lead(datetime),datetime))) |> 
  mutate(steplength_m = sqrt((lon_x-lead(lon_x))^2+ (lat_y-lead(lat_y))^2)) |> 
  mutate(speed_ms = steplength_m/diff_s)

hist(log10(posmo_neu$speed_ms))


# plot the data using tmap
tm_shape(posmo_neu) +
   tm_dots(col = "red", size = 0.2) +
  tm_basemap(server = "OpenStreetMap")


# download a map of Zurich from Google Maps
zurich_map <- qmap("Zurich", zoom = 13, maptype = "terrain")

# convert the map to an sf object
zurich_sf <- ggmap::ggmap_to_sf(zurich_map)

# plot the data and map using tmap
tm_shape(zurich_sf) +
  tm_raster() +
  tm_shape(tracking_sf) +
  tm_dots(col = "red", size = 1) +
  tm_basemap(server = "OpenStreetMap")
```

